---
title: "Wine Quality Prediction Project"
authors: "Autumn Heyman, Erin Weaver, Georgia Miller"
output: html_document
---
# File Setup Tasks
## Import Libraries

```{r}
library("tidyverse")
library("caret")
library("lmtest")
library("magrittr")
library("dplyr")
library("tidyr")
library("popbio")
library("e1071")
library("PerformanceAnalytics")
library("corrplot")
library("corpcor")
library("MASS")
library("brant")
library("car")
library("effects")
```

## Set Working Directory
*** can be manually overridden by selecting source file location from the session menu **

```{r}
setwd("/Users/erinweaver/Documents/GitHub/TheThreeMusketeers")
```

## Import Dataset

```{r}
wineqt <- read.csv("/Users/erinweaver/Documents/GitHub/TheThreeMusketeers/Data/WineQT.csv")
head(wineqt)
```

# Data Wrangling 

## Drop Id Column 

```{r}
wine_subset <- wineqt[, c(1,2,3,4,5,6,7,8,9,10,11,12)]
```

## Recode into three groups

```{r}

wine_subset$qualityR <- NA
wine_subset$qualityR [wine_subset$quality==3] <- 0
wine_subset$qualityR [wine_subset$quality==4] <- 0
wine_subset$qualityR [wine_subset$quality==5] <- 1
wine_subset$qualityR [wine_subset$quality==6] <- 1
wine_subset$qualityR [wine_subset$quality==7] <- 2
wine_subset$qualityR [wine_subset$quality==8] <- 2
head(wine_subset)

```
## Drop Quality

```{r}
wine_wrangled <- wine_subset[, c(1,2,3,4,5,6,7,8,9,10,11,13)]
head(wine_wrangled)
```

## Check All Data Types 

```{r}
str(wine_wrangled)
```


# Exploratory Analysis

### Summarizing the dataset
```{r}
summary(wine_wrangled)
```

********* STOP SIGN ******** :)

# Assumption Testing:

## Assumption 1: The Dependent Variable is Ordered
### Results of Test: Pass.

## Assumption 2: One or more of the independent variables are either continuous, categorical, or ordinal
### Results of Test: Pass.

## Assumption 3: Absence of Multicollinearity

### Correlation Matrix
```{r}
wine_wrangled$qualityR <- as.numeric(wine_wrangled$qualityR)
corr_matrix <- cor(wine_wrangled)
corr_matrix
```

### Correlation Plots

```{r}
corrplot(corr_matrix, type="upper", order="hclust", p.mat = corr_matrix, sig.level = 0.01, insig="blank")
```

```{r}
corrplot(corr_matrix, method = "number", is.corr = FALSE)
```

### Pearson's Test

```{r}
chart.Correlation(wine_wrangled, histogram=FALSE, method="pearson")
```
### Results of Test: Fail, assuming the cut off is 0.6 for correlation, Pass if the cut off is 0.7 for correlation.

The Variable Pairs with the highest correlation are as follows:
* pH & fixed.acidity (-0.69)
* density & fixed.acidity (0.68)
* citric.acid & fixed.acidity (0.67)
* total sulfur.dioxide & free.sulfur.dioxide (0.66)
--> should we drop fixed.acidity & free.sulfur.dioxide?


# Creating a Base Model  

## Define the Order of the Dependent Variable
```{r}

wine_wrangled$qualityR = factor(wine_wrangled$qualityR, levels = c("0", "1", "2"), ordered = TRUE) 
str(wine_wrangled)

```

## Dividing data into training and test set  *** REVIEW THIS SECTION ***
*** how to determine appropriate sample size
*** what is seed and how to determine appropriate value

### Random sampling 
```{r}

samplesize = 0.60*nrow(wine_wrangled)
set.seed(100)
index = sample(seq_len(nrow(wine_wrangled)), size = samplesize)

```

### Creating training and test set 
```{r}
datatrain = wine_wrangled[index,]
datatest = wine_wrangled[-index,]
```

## Create the Ordinal Logistic Regression Model
```{r}

model= polr(qualityR ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = wine_wrangled, Hess = TRUE)

summary(model)

```

## Significance Table of coefficients and intercepts

```{R}
summary_table <- coef(summary(model))
pval <- pnorm(abs(summary_table[, "t value"]),lower.tail = FALSE)* 2
summary_table <- cbind(summary_table, "p value" = round(pval,3))
summary_table
```

ordinaloutput <- as.data.frame(summary_table)
ordinaloutput

library(tibble)
ordinaloutput <- tibble::rownames_to_column(ordinaloutput, "Variable")

write.csv(ordinaloutput,"/Users/erinweaver/Documents/GitHub/TheThreeMusketeers/Data/ordinaloutput.csv", row.names = FALSE)


## Explanation of Results:


## Assumption 4: Proportional Odds  - run after model is generated

```{r}
brant(model)
```

### Results of Test: Pass


# Making predictions

## Compute Confusion Matrix and Misclassification Error 

```{r}
predict.quality = predict(model,datatest)
table(datatest$qualityR, predict.quality)
mean(as.character(datatest$qualityR) != as.character(predict.quality))
```

### Interpretation of the Confusion Matrix
###In the test dataset, 0 times the below average "0" wine quality category is identified correctly. We observe that the model identifies below average wine poorly. This is because there is an inadequate representation of "below average" wine quality category in the traning dataset.
###In the test dataset, 352 times the average "1" wine quality category is identified correctly.
###In the test dataset, 23 times the above average "2" wine quality category is identified correctly.
###Using the confusion matrix, we find that the misclassification error for our model is 18%.


## Interpretation Using Plots: Plotting the Effects

### Fixed Acidity
```{r}
Effect(focal.predictors = "fixed.acidity",model)
plot(Effect(focal.predictors = "fixed.acidity",model))
```

### Volatile Acidity
```{r}
Effect(focal.predictors = "volatile.acidity",model)
plot(Effect(focal.predictors = "volatile.acidity",model))
```

### Citric Acid
```{r}
Effect(focal.predictors = "citric.acid",model)
plot(Effect(focal.predictors = "citric.acid",model))
```

### Residual Sugar
```{r}
Effect(focal.predictors = "residual.sugar",model)
plot(Effect(focal.predictors = "residual.sugar",model))
```

### Chlorides
```{r}
Effect(focal.predictors = "chlorides",model)
plot(Effect(focal.predictors = "chlorides",model))
```

### Free Sulfur Dioxide
```{r}
Effect(focal.predictors = "free.sulfur.dioxide",model)
plot(Effect(focal.predictors = "free.sulfur.dioxide",model))
```

### Total Sulfur Dioxide
```{r}
Effect(focal.predictors = "total.sulfur.dioxide",model)
plot(Effect(focal.predictors = "total.sulfur.dioxide",model))
```

### Density
```{r}
Effect(focal.predictors = "density",model)
plot(Effect(focal.predictors = "density",model))
```

### pH
```{r}
Effect(focal.predictors = "pH",model)
plot(Effect(focal.predictors = "fixed.acidity",model))
```

### Sulphates
```{r}
Effect(focal.predictors = "sulphates",model)
plot(Effect(focal.predictors = "sulphates",model))
```

### Alcohol
```{r}
Effect(focal.predictors = "alcohol",model)
plot(Effect(focal.predictors = "alcohol",model))
```












**** SECTIONS OF THE CODE BELOW IS TEMPORARILY DISABLED BY HASH TAGS ****

## Assumption #5: Logit Linearity *** MODIFY TO FIT NEW ANALYSIS ***

### create probabilities
```{r}
#prob <- predict(model, type="probs")
#head(prob)
```

### define predictors
```{r}
#wine1 <- wine_wrangled %>% dplyr::select_if(is.numeric)
#predictors <- colnames(wine1)
```

### 
```{r}
#wine1 <- wine1 %>% mutate(logit=log(prob/(1-prob))) %>%
#gather(key= "predictors", value="predictor.value", -logit)
```

## Assumption #5: Absence of Influential Outliers *** MODIFY TO FIT NEW ANALYSIS ***

```{r}
#infl <- influence.measures(model)
#summary(infl)
```
### Results indicate no evidence of **influential** outliers

Box Plot Method

### all box plots
```{r}
boxplot(wine_wrangled)
ggplot(stack(wine_wrangled), aes(x = ind, y = values))+
  geom_boxplot(fill='rosybrown', color="darkred") +
  coord_flip()
```

All variables have outliers.

*********************** MOVING FORWARD **************************

# Correcting for Violation of Assumptions

## Corrections for Violation of Multicollinearity

```{r}
```

## Corrections for Violation of Logit Linearity
```{r}
```


